What can GitHub and Googlecode Issues tell us about Exceptions in Android?

### 2014-08-23 android-stacks.tex
changing the focus. More inprovements in sec 2

\title{ What can GitHub and Googlecode Issues tell us about Exceptions in Android Apps? }

@@

\subsection{Java Exception Handling Mechanism} \label{sec:extypes}

In Java, an exception is represented according to the class hierarchy shown in
Figure~\ref{fig:exchier}.  According to it every exception is an
instance of the Throwable class, and can be of three kinds: the checked exceptions
(extends Exception) - means that the compiler statically checks if appropriate handlers 
are provided within the system, to enable such checking each checked exception must be specified 
on the method signature;  the run-time exceptions (extends RuntimeException) and errors
(extends Error)~\cite{gosling2000java}  -  which do not need to be specified on method
signatures, and hence both are referred as unchecked exceptions.

The Java Virtual Machine represents as runtime exceptions invalid operations detected in
the program (out-of-bounds array index, divide-by-zero error, null pointer
references) from which most of the programs are not expected to recover. By
convention an Error represents an unrecoverable condition which usually results
from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
normally should not be handled inside the application. The Java specification~\cite{gosling2000java} 
suggests that user-defined exceptions should be checked exceptions. The main reason is because doing so the
callers of a method will know about the exceptions that a method can throw, and
so they can decide what to do about them.

Hence, in Java an exception can be thrown in one of the following
circumstances~\cite{gosling2000java}: 
(i) explicitly thrown when a throw statement is executed; 
(ii) implicitly thrown by the JVM when the evaluation of an expression
 violates the normal semantics of language, also refered as a coding error
 (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
(iii) implicitly thrown by the JVM due to an internal error or resource limitation (e.g.,
OutofMemoryError).

%\todo{This sub-section is quite similar to the section introduction, shouldn't
%we merge them?}

When an exception is throw, it causes a transfer of control from the point
where the exception occurred to a point that can be specified by the programmer
(exception handler) - in Java it is represented by the try-catch block . A common way of  
propagating exceptions in Java programs is through exception wrapping
 (also known as exception chaining). Exception wrapping allows one exception to be wrapped in another one and re-thrown. 
Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
exception wrapping. The bottom part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

\begin{figure} \centering \includegraphics[scale=0.]{stacktrace.png}
\caption{Example of an Exception stack trace in Java.}
\label{fig:wrapping}
\end{figure}

@@

@@

\subsection{Meaning of Exception Types and related Best Practices}
\label{sec:best}

Several general guidelines have been proposed on how to use Java
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective}. 
We compiled a list of the guidelines related to exception types and exception
propagation in Java the following:\footnote{We also compiled guidelines related
to exception handling but they are out of the scope of this paper.}

\textbf{I-Use checked exceptions to represent recoverable
conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective})
The developer should use checked exceptions for conditions from which the caller
is expected to recover. By confronting the API user with a checked exception,
the API designer is forcing the client to handle the exceptional condition. The
client can explicitly ignore the exception (swallowing, or converting it to
other type) at the expense of the program's robustness~\cite{gosling2000java}.

\textbf{III-A method should throw exceptions that precisely define the
exceptional condition} (\cite{gosling2000java,bloch2008effective}). To do so,
developers should either try to reuse the exception types already defined in the
JVM or they should create a specific exception. Throwing general types such as a
pure Exception or a RuntimeException is hence considered a bad practice.

%\noindent\emph{Exception Documentation}
\textbf{ IV-Document all exceptions thrown by a
method} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled JVM runtime exceptions related to programming
mistakes} as carefully as checked exceptions. Doing so, the clients of a method
will be aware of all the exceptions the method can throw. If the developer fails to
do follow this practice (specially when developing library code) it will be
difficult or even impossible for the caller to make effective use of such 
method~\cite{wirfs2006toward, bloch2008effective}. As mentioned in Section II, 
in some languages such as C

@@

\subsection{Investigating The Android Subset}

@@

\begin{table*}
  \centering
  \begin{tabular}{rccccccc}
    \hline
    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{\#Projects}  &  \bfseries{\#Occurrences} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib}  \\
    \hline

%\begin{table}
%\centering
%\begin{tabular}{llll}
%    \hline
%    \bfseries{Wrapper Type} & \bfseries{Root Cause Type} & \bfseries{Occurrences} & \bfseries{Projects} \\
%    \hline
      %none   & Runtime    & 359 &  2360  & - &  - \\
      %none  &   Checked   & 117 & 422 & - &  -   \\
      %none  &   Error    & 141 & 381 & - &  -  \\
      %none &  Undefined  & 10  &  15  & - &  -    \\
      %none & Throwable   & 1  &  1 & - &  -   \\
     %\hline
      %Runtime   & Runtime  & 182 & 560 & - &  -  \\    
      %Checked   & Checked   & 42 & 98  & - &  -  \\
      %Error     & Error    & 14 & 15  & - &  -   \\
    %\hline

     Runtime &  Checked   & 70 & 114 & 0 &  0  & 0 &  0   \\
      Runtime   &  Error   & 38  & 57 & 0 &  0  & 0 &  0       \\
      Checked &  Runtime   & 12 & 22 & 0 &  0  & 0 &  0        \\
      Checked & Error      & 7 &  8 & 0 &  0  & 0 &  0       \\
      Error & Checked      & 4  &  7 & 0 &  0  & 0 &  0     \\
      Error & Runtime     & 6  &  13   & 0 &  0  & 0 &  0    \
    % Throwable & Runtime   & 1 &  3   & 0 &  0  & 0 &  0    \\
    \hline
  \end{tabular}
\caption{Kinds of wrappings found on the stacks of 482 ANDROID projects found on Github}
\label{tab:wrappingandroid}
\end{table*}

@@

%\subsection{Actionable Advices}
%The results of a mining study should be a set of actionalble advices (tao xie).

>>>>

\title{ What can GitHub and Googlecode Issues tell us about Exceptions in Android? }

@@

@@

- opensource
- elements of archtecture.
- Inversion of control development model.
- Activity onStart onResume onStop...
- Acess and share resources...

\subsection{Exceptions in Java} \label{sec:extypes}

In Java, an exception is represented according to the class hierarchy shown in
Figure~\ref{fig:exchier}.  According to it every exception is an
instance of the Throwable class, and can be of three kinds: the checked exceptions
(extends Exception), the runtime exceptions (extends RuntimeException) and errors
(extends Error)~\cite{gosling2000java}.

The checked exceptions must be specified on the method signature that may throw it, 
and the compiler statically checks if appropriate handlers are provided within the system.
Both runtime exceptions and errors are known as unchecked exceptions, because 
they do not need to be specified on method signatures and do not trigger any 
compile time checking.

By convention an Error represents an unrecoverable condition which usually results
from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
normally should not be handled inside the application. The user-defined exceptions 
can be either checked or runtime. Although the Java specification~\cite{gosling2000java} 
suggests that user-defined exceptions should be checked exception (because doing so 
the callers of a method will know about the exceptions that a method can throw and so 
they can decide what to do about them) there is a long lasting debate about the use of 
checked and unchecked exceptions in Java.

The runtime exceptions, however are not only used to represent user-defined
exceptional conditions, they are used by the JVM when a program violates 
the semantic constraints of the Java programming language (out-of-bounds array index, divide-by-zero 
error, null pointer references) from which most of the programs are not expected to recover. 
\footnote{Some programming languages react to such errors by peremptorily terminating the program; 
other programming languages allow similar approaches e.g.C\#; and others let the program to continue
 its execution in some situations such as the out-of-bounds array index e.g., C++. }

%In Java an exception can be thrown in one of the following
%circumstances~\cite{gosling2000java}: 
%(i) explicitly thrown when a throw statement is executed; 
%(ii) implicitly thrown by the JVM when the evaluation of an expression
% violates the normal semantics of language, also refered as a coding error
% (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
%(iii) implicitly thrown by the JVM due to an internal error or resource limitation (e.g.,
%OutofMemoryError). 
%\todo{This sub-section is quite similar to the section introduction, shouldn't
%we merge them?}

When an exception is thrown, it causes a transfer of control from the point
where the exception occurred to a point that can be specified by the programmer
(exception handler) - in Java it is represented by the try-catch block . A common way of  
propagating exceptions in Java programs is through exception wrapping
 (also known as exception chaining). Exception wrapping allows one exception 
to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} presents 
a stack trace associated which illustrates an exception wrapping. The bottom 
part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

\begin{figure} \centering \includegraphics[scale=0.8]{stacktrace.png}
\caption{Example of an Exception stack trace in Java.}
\label{fig:wrapping}
\end{figure}

\subsection{Related Best Practices}
\label{sec:best}

Several general guidelines have been proposed on how to use Java
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective}. 
We compiled a list of the guidelines related to exception types and exception propagation in Java, summarized below:\footnote{We also compiled guidelines related
to exception handling but they are out of the scope of this paper.}

\textbf{I-Checked exceptions should be used to represent recoverable
conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective})
The developer should use checked exceptions for conditions from which the caller
is expected to recover. By confronting the API user with a checked exception,
the API designer is forcing the client to handle the exceptional condition. The
client can explicitly ignore the exception (swallowing, or converting it to
other type) at the expense of the program's robustness~\cite{gosling2000java}.

\textbf{III-A method should throw exceptions that precisely define the
exceptional condition} (\cite{gosling2000java,bloch2008effective}). To do so,
developers should either try to reuse the exception types already defined in the
JVM or they should create a specific exception. Throwing general types such as a
pure java.lang.Exception or a java.lang.RuntimeException is considered a bad practice.

%\noindent\emph{Exception Documentation}
%\textbf{IV- It is wise to document all exceptions thrown by a method}

\textbf{IV- All exceptions explicitly thrown by a method should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled JVM runtime exceptions related to programming
mistakes} as carefully as checked exceptions. Doing so, the clients of a method
will be aware of all the exceptions the method can throw. If the developer fails to
do follow this practice (specially when developing library code) it will be
difficult or even impossible for the caller to make effective use of such 
method~\cite{wirfs2006toward, bloch2008effective} and protect its code
against unforeseen exceptions.

@@

@@

@@

\begin{table*}
  \centering
  \begin{tabular}{llcccccc}
    \hline
    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{Projects}  &  \bfseries{Occurrences} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib}  \\
    \hline
      Runtime &  Checked   & 70 & 114 & 0 &  0  & 0 &  0 \\
      Runtime   &  Error   & 38  & 57 & 0 &  0  & 0 &  0 \\
      Checked &  Runtime   & 12 & 22 & 0 &  0  & 0 &  0 \\
      Checked & Error      & 7 &  8 & 0 &  0  & 0 &  0 \\
      Error & Checked      & 4  &  7 & 0 &  0  & 0 &  0     \\
      Error & Runtime     & 6  &  13   & 0 &  0  & 0 &  0    \\
     Throwable & Runtime   & 1 &  3   & 0 &  0  & 0 &  0    \\
    \hline
  \end{tabular}
\caption{Wrappings comprising different exception types.}
\label{tab:wrappingandroid}
\end{table*}

@@

\subsection{Actionable Advices - The Unwrittern rules of Android development}

%The results of a mining study should be a set of actionalble advices (tao xie).
